### 2019-09-27 paper/8.related-work.tex

\section{Related Work}

\begin{itemize}
    %\item BiGUL papers (PEPM, POPL, tutorial?)
    %\item the first BX paper
    %\item get-based BX
    %\item put-based BX?
    \item implementation of BX, they might have the same problem, this paper will be first attempt, semantics and correctness is considered more, but efficiency is  lacked.
    \item optimization techniques : tupling, lazy computation, fusion transformation?
    \item (view update problem paper)
\end{itemize}