### 2019-11-15 paper/0.abst.tex
extract abstract from main.tex, update abstract

 Bidirectional transformation (BX) is a solution to the view update problem and widely used for synchronizing data. Its semantics and correctness are researched well, but the evaluation efficiency is not considered so far. As a first attempt we consider efficiency of evaluation in this work. To improve evaluation efficiency, we focus on the evaluation of BX compositions, because its evaluation often causes redundant evaluations. 
 For evaluating BX compositions efficiently, we introduce two memoization methods. The first method, minBigul\_m, uses simple memoization, but it works well for many BX programs. This keeps the evaluated results in a table and reuses them.  A disadvantage of this is cost for saving to a table and searching. When the inputs are large, the evaluation time becomes slow. To treat large inputs, we introduce the second method, xpg, that uses tupling, lazy update and lazy evaluation as optimizations. Lazy update keeps updates in closures and enables to use them later.
 The experiments of these methods show the following results: the original method that is used in the language is faster for right associative compositions, minBiGUL\_m is faster for not-large inputs and xpg is faster for large inputs.