### 2020-02-21 paper/main.tex
removed unneeded part in main.tex

% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.20 of 2017/10/04
%
\documentclass[runningheads]{llncs}
%
\usepackage{graphicx}
\usepackage{color}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{multirow}
\usepackage{multicol}
\usepackage{tabularx}
\usepackage{tikz}
\usepackage{pgfplots}
\usepackage{pgfplotstable}
\usepackage{lmodern}
\usepackage{pgffor}
\usepackage{schemata}
\usepackage[colorlinks=true,linkcolor=blue,citecolor=blue]{hyperref}

\usepackage[backend=bibtex,doi=false,url=false,style=nature,sorting=none]{biblatex}
\addbibresource{bibliography}

% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following line
% to display URLs in blue roman font according to Springer's eBook style:
\renewcommand\UrlFont{\color{blue}\rmfamily}
\renewcommand{\footnotesize}{\scriptsize}

\usetikzlibrary{matrix}
\usepgfplotslibrary{groupplots}
\pgfplotsset{compat=newest}

% \setlength{\parindent}{0pt}
% \renewcommand{\indent}{\hspace*{0pt}}

\newcommand{\tab}{\hspace*{3mm}}
\newcommand{\tabs}[1]{\foreach \n in {1,...,#1}{\tab}}
\newcommand{\qtab}{\hspace*{3mm} \ \quad}
\newcommand{\qtabs}[1]{\foreach \n in {1,...,#1}{\qtab}}

\newcommand{\sif}[3]{\textnormal{if} \ #1 \ \textnormal{then} \ #2 \ \textnormal{else} \ #3}
\newcommand{\product}[2]{#1 \times #2}
\newcommand{\tuple}[2]{(#1 :: #2)}
\newcommand{\rearrs}[3]{RearrS \ #1 \ #2 \ #3}
\newcommand{\rearrv}[3]{RearrV \ #1 \ #2 \ #3}
\newcommand{\casebx}[4]{Case \ #1 \ #2 \ #3 \ #4}

\newcommand{\putbx}[3]{put \, [\![#1]\!] \ #2 \ #3}
\newcommand{\putbxinline}[1]{put \, [\![#1]\!]}
\newcommand{\getbx}[2]{get \, [\![#1]\!] \ #2}
\newcommand{\getbxinline}[1]{get \, [\![#1]\!]}

\newcommand{\putm}[3]{put_m \, [\![#1]\!] \ #2 \ #3}
\newcommand{\putminline}[1]{put_m \, [\![#1]\!]}
\newcommand{\getm}[2]{get_m \, [\![#1]\!] \ #2}
\newcommand{\getminline}[1]{get_m \, [\![#1]\!]}

\newcommand{\putrev}[3]{put_{REV} \, [\![#1]\!] \ {#2} \ {#3}}
\newcommand{\getrev}[2]{get_{REV} \, [\![#1]\!] \ {#2}}

\newcommand{\pg}[3]{pg \, [\![#1]\!] (#2, #3)}
\newcommand{\pginline}[1]{pg \, [\![#1]\!]}

\newcommand{\cpg}[5]{cpg [\![#1]\!] (#2, #3, #4, #5)}
\newcommand{\cpginline}[1]{cpg \, [\![#1]\!]}

\newcommand{\kpg}[7]{kpg [\![#1]\!] (#2, #3, #4, #5, #6, #7)}
\newcommand{\kpginline}[1]{kpg \, [\![#1]\!]}

\newcommand{\xpg}[3]{xpg \, [\![#1]\!] (#2, #3)}
\newcommand{\xpginline}[1]{xpg \, [\![#1]\!]}
\newcommand{\smallvspace}{\vspace{1.2mm}}
\newcommand{\minusvspace}{\vspace*{-5.5mm}}
\newcommand{\minusvspacetwo}{\vspace*{-3.5mm}}

\newcommand{\match}[1]{\textnormal{match } #1 \ \textnormal{ with }}

\newcommand{\newcirc}{\tilde{\circ}}

\makeatletter
\newcommand{\AlignFootnote}[1]{%
    \ifmeasuring@
    \else
        \footnote{#1}%
    \fi
}
\makeatother

\newcolumntype{L}[1]{>{\raggedright\arraybackslash}p{#1}}
\newcolumntype{C}[1]{>{\centering\arraybackslash}p{#1}}
\newcolumntype{R}[1]{>{\raggedleft\arraybackslash}p{#1}}

\begin{document}
%
\title{
  % Strategies for efficient composition of bidirectional programs
  % by memoization and lazy updates
  %  tupling
  %  accumulating updates = lazy updates
  %  based on the idea of reversible computing  
  % \thanks{Supported by organization x.}  
  An Efficient Composition of Bidirectional Programs
  by Memoization and Lazy Update
}
%
\titlerunning{An Efficient Composition of BX Programs}
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%

\author{{Kanae Tsushima\inst{1}} \and %\break% \orcidID{0000-0002-3383-3389}
  {Bach Nguyen Trong\inst{1}} \and %\break %\orcidID{0000-0002-6342-6020}
  {Robert Gl\"{u}ck\inst{2}} \and %\break %\orcidID{0000-0001-6990-3935}
  {Zhenjiang Hu\inst{3}}} %\orcidID{0000-0002-9034-205X}

%\author{anonymous authors}

\authorrunning{K. Tsushima et al.}

%\authorrunning{anonymous authors}

% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%

\institute{National Institute of Informatics, Japan \break
  \email{\{k\_tsushima,bach\}@nii.ac.jp} \and
  University of Copenhagen, Denmark \break
  \email{glueck@acm.org} \and
  Peking University, China \break
  \email{huzj@pku.edu.cn}}

%\institute{anonymous}

\pagestyle{empty}

%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
%The abstract should briefly summarize the contents of the paper in
%15--250 words.
 \input{0.abst.tex}
 % We show experimental results for several BX programs with simple inputs and large inputs.

  %To achieve fast evaluation for complicated data, we introduced a novel approach. First, we make $put$ and $get$ more tight, based on the idea of tupling. Because simple tupled result includes redundancies for left associative compositions, we apply two optimization techniques: lazy update and lazy evaluation. We compare the approaches including the strategy used in the actual BX language, and show that our optimized approach is faster than the other approaches (?).

\keywords{Bidirectional transformation \and
  Implementation technique \and
  Efficiency \and 
  Optimization \and Tupling.
}
\end{abstract}

\input{1.introduction.tex}

\input{2.minbigul.tex}

\input{3.minbigul_m.tex}

\input{4.xpg.tex}

\input{5.experiments.tex}

\input{6.related-work.tex}

\input{7.conclusion.tex}

\subsubsection*{Acknowledgement}
We thank the anonymous reviewers for valuable comments and suggestions. This work has been partially supported by JSPS KAKENHI Grant Number JP17H06099.

% From sample code
% 
% ---- Bibliography ----
%
% BibTeX users should specify bibliography style 'splncs04'.
% References will then be sorted and formatted in the correct style.
%
% \bibliography{bibliography}
% \bibliographystyle{splncs04}

\printbibliography

\end{document}